# Look at MIC levels for each AM:

### TET- Tetracyclines

```{r}
table(NARMSCattleCombined_ordinal$TET,NARMSCattleCombined_ordinal$TET.Sign)
#will not merge any categories
```

#NARMS breakpoints given by FDA

-   S: \<= 4

-   I: 8

-   **R: \>= 16**

#ECOFF Value from EUCAST: **8**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

not needed

### NAL: Quinolones

```{r}
table(NARMSCattleCombined_ordinal$NAL,NARMSCattleCombined_ordinal$NAL.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 16

-   I: N/A

-   **R: \>= 32**

#ECOFF Value from EUCAST: **8**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

New Categories:

-   \<=1

-   =2

-   8

    -   sign \>2-\<=8

    -   combined =8 with =4

-   =16

-   \>=32

```{r}
NARMSCattleCombined_ordinal$NAL[NARMSCattleCombined_ordinal$NAL==0.5]<- 1
NARMSCattleCombined_ordinal$NAL[NARMSCattleCombined_ordinal$NAL==4]<- 8

table(NARMSCattleCombined_ordinal$NAL)
NARMSCattleCombined_ordinal$NAL<- droplevels(NARMSCattleCombined_ordinal$NAL)
table(NARMSCattleCombined_ordinal$NAL)
```

### CIP: Quinolones

```{r}
table(NARMSCattleCombined_ordinal$CIP,NARMSCattleCombined_ordinal$CIP.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 0.06

-   I: \>= 0.12

-   **R: \>= 0.12**

#ECOFF Value from **EUCAST: 0.06**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels

New Categories:

-   \<= 0.015

-   0.06

    -   sign is \>0.015- \<=0.06

    -   combine 0.03 and 0.06 categories

-   \>= 0.125

```{r}
NARMSCattleCombined_ordinal$CIP[NARMSCattleCombined_ordinal$CIP==0.03]<- 0.06
NARMSCattleCombined_ordinal$CIP[NARMSCattleCombined_ordinal$CIP==0.25]<- 0.125
NARMSCattleCombined_ordinal$CIP[NARMSCattleCombined_ordinal$CIP==0.5]<- 0.125
NARMSCattleCombined_ordinal$CIP[NARMSCattleCombined_ordinal$CIP==1]<- 0.125
NARMSCattleCombined_ordinal$CIP[NARMSCattleCombined_ordinal$CIP==2]<- 0.125
NARMSCattleCombined_ordinal$CIP[NARMSCattleCombined_ordinal$CIP==4]<- 0.125

table(NARMSCattleCombined_ordinal$CIP)
NARMSCattleCombined_ordinal$CIP<- droplevels(NARMSCattleCombined_ordinal$CIP)
table(NARMSCattleCombined_ordinal$CIP)
```

### CHL: Phenicol

```{r}
table(NARMSCattleCombined_ordinal$CHL,NARMSCattleCombined_ordinal$CHL.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 8

-   I: \>= 16

-   **R: \>=32**

#ECOFF Value from **EUCAST: 16**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

New categories

-   \<= 4

-   =8

-   =16

-   \>=32

```{r}
NARMSCattleCombined_ordinal$CHL[NARMSCattleCombined_ordinal$CHL==2]<- 4

table(NARMSCattleCombined_ordinal$CHL)
NARMSCattleCombined_ordinal$CHL<- droplevels(NARMSCattleCombined_ordinal$CHL)
table(NARMSCattleCombined_ordinal$CHL)
```

### AMP: Phenicol

```{r}
table(NARMSCattleCombined_ordinal$AMP, NARMSCattleCombined_ordinal$AMP.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 8

-   I: \>= 16

-   **R: \>=32**

#ECOFF Value from **EUCAST: 8**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

will not change

### AZI: Macrolide

```{r}
table(NARMSCattleCombined_ordinal$AZI, NARMSCattleCombined_ordinal$AZI.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<=16

-   I: \>= N/A

-   **R: \>=32**

#ECOFF Value from **EUCAST: 16**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels

New categories:

-   \<=1

-   =2

-   =4

-   =8

-   =16

-   \>=32

```{r}
NARMSCattleCombined_ordinal$AZI[NARMSCattleCombined_ordinal$AZI==0.25]<- 1
NARMSCattleCombined_ordinal$AZI[NARMSCattleCombined_ordinal$AZI==0.5]<- 1

table(NARMSCattleCombined_ordinal$AZI)
NARMSCattleCombined_ordinal$AZI<- droplevels(NARMSCattleCombined_ordinal$AZI)
table(NARMSCattleCombined_ordinal$AZI)
```

### COT: Folate Pathway Inhibitor

```{r}
table(NARMSCattleCombined_ordinal$COT, NARMSCattleCombined_ordinal$COT.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 2

-   I: \>= N/A

-   **R: \>=4**

#ECOFF Value from **EUCAST: 0.5**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

New Categories:

-   \<=0.125

-   0.5

    -   sign is \>0.125- \<=0.5

    -   combine 0.25 and 0.5

-   2

    -   sign is \>0.5- \<=2

    -   combine 1 and 2

-   \>=4

```{r}
NARMSCattleCombined_ordinal$COT[NARMSCattleCombined_ordinal$COT==0.25]<- 0.5
NARMSCattleCombined_ordinal$COT[NARMSCattleCombined_ordinal$COT==1]<- 2

table(NARMSCattleCombined_ordinal$COT)
NARMSCattleCombined_ordinal$COT<- droplevels(NARMSCattleCombined_ordinal$COT)
table(NARMSCattleCombined_ordinal$COT)
```

### FIS (Folate Pathway Inhibitor)

```{r}
table(NARMSCattleCombined_ordinal$FIS, NARMSCattleCombined_ordinal$FIS.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 256

-   I: \>= NA

-   **R: \>=512**

#ECOFF Value from **EUCAST: No data available**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

New categories

-   \<=16

-   =32

-   128

    -   sign is \>32- \<=128

    -   combine categories 64 and 128

-   \>=256

```{r}
NARMSCattleCombined_ordinal$FIS[NARMSCattleCombined_ordinal$FIS==64]<- 128

table(NARMSCattleCombined_ordinal$FIS)
NARMSCattleCombined_ordinal$FIS<- droplevels(NARMSCattleCombined_ordinal$FIS)
table(NARMSCattleCombined_ordinal$FIS)
```

### AXO (Cephem)

```{r}
table(NARMSCattleCombined_ordinal$AXO,NARMSCattleCombined_ordinal$AXO.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 1

-   I: \>= 2

-   **R: \>=4**

#ECOFF Value from **EUCAST: 0.125**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

it's so unbalanced, will skip making a model for AXO

### FOX (Cephem)

```{r}
table(NARMSCattleCombined_ordinal$FOX,NARMSCattleCombined_ordinal$FOX.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 8

-   I: \>= 16

-   **R: \>=32**

#ECOFF Value from **EUCAST: 16**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

New Categories:

-   \<=2

    -   combine categories 0.5, 1, and 2

-   =4

-   =8

-   =16

-   \>=32

```{r}
NARMSCattleCombined_ordinal$FOX[NARMSCattleCombined_ordinal$FOX==0.5]<- 2
NARMSCattleCombined_ordinal$FOX[NARMSCattleCombined_ordinal$FOX==1]<- 2

table(NARMSCattleCombined_ordinal$FOX)
NARMSCattleCombined_ordinal$FOX<- droplevels(NARMSCattleCombined_ordinal$FOX)
table(NARMSCattleCombined_ordinal$FOX)
```

### TIO (Cephem)

```{r}
table(NARMSCattleCombined_ordinal$TIO, NARMSCattleCombined_ordinal$TIO.Sign)
```

#NARMS breakpoints given by USDA (not on FDA list)

-   S: \<= 2

-   I: \>=4

-   **R: \>=8**

#ECOFF Value from **EUCAST: 1**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

Due to not wanting to cross NARMS or ECOFF boundaries, will not change categories

### AMC (Beta lactam)

```{r}
table(NARMSCattleCombined_ordinal$AMC,NARMSCattleCombined_ordinal$AMC.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 8

-   I: \>= 16

-   **R: \>=32**

#ECOFF Value from **EUCAST: 8**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

Due to not wanting to cross NARMS or ECOFF boundaries, will not change categories

### Str (aminoglycoside)

```{r}
table(NARMSCattleCombined_ordinal$Str, NARMSCattleCombined_ordinal$STR.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 16

-   I: \>= NA

-   **R: \>=32**

#ECOFF Value from **EUCAST: 16**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

New Categories:

-   \<=4

    -   combine categories 2 and 4

-   =8

-   =16

-   \>=32

```{r}
NARMSCattleCombined_ordinal$Str[NARMSCattleCombined_ordinal$Str==2]<- 4

table(NARMSCattleCombined_ordinal$Str)
NARMSCattleCombined_ordinal$Str<- droplevels(NARMSCattleCombined_ordinal$Str)
table(NARMSCattleCombined_ordinal$Str)
```

### KAN (aminoglycoside)

```{r}
table(NARMSCattleCombined_ordinal$KAN, NARMSCattleCombined_ordinal$KAN.Sign)
```

#NARMS breakpoints given by USDA (FDA not available)

-   S: \<= 16

-   I: \>= 32

-   **R: \>=64**

#ECOFF Value from **EUCAST: 16**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

Due to not wanting to cross NARMS or ECOFF boundaries, will not change categories

### GEN (aminoglycoside)

```{r}
table(NARMSCattleCombined_ordinal$GEN, NARMSCattleCombined_ordinal$GEN.Sign)
```

#NARMS breakpoints given by FDA

-   S: \<= 4

-   I: \>= 8

-   **R: \>=16**

#ECOFF Value from **EUCAST: 2**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

Due to not wanting to cross NARMS or ECOFF boundaries, will not change categories

### AMI (aminoglycoside)

```{r}
table(NARMSCattleCombined_ordinal$AMI,NARMSCattleCombined_ordinal$AMI.Sign)
```

#NARMS breakpoints given by USDA

-   S: \<= 16

-   I: \>= 32

-   **R: \>=64**

#ECOFF Value from **EUCAST: 8**

-   Isolates of a species that have MIC values at or below the ECOFF are members of the wild-type population for the species and agent in question

##### Condense Levels:

New Categories:

-   \<=1

    -   combine categories 0.5 and 1

-   =2

-   =4

-   =8

-   \>=16

    -   will not combine with 8 due to ECOFF boundary

```{r}
NARMSCattleCombined_ordinal$AMI[NARMSCattleCombined_ordinal$AMI==0.5]<- 1

table(NARMSCattleCombined_ordinal$AMI)
NARMSCattleCombined_ordinal$AMI<- droplevels(NARMSCattleCombined_ordinal$AMI)
table(NARMSCattleCombined_ordinal$AMI)
```
