### Look at +/-2 Category Accuracies

MIC data distributions can have a =/-2 dilutions margin of error (why quality control ranges for susceptibility testing machines are so big)

##### AMC

```{r}
PlusorMinus2_overall(final_AMC_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_AMC_RF_ConMat)
```

| **MIC isolate category** | **Proportion_of_Isolates** |
|:-------------------------|:---------------------------|

|     |          |
|-----|----------|
| 1   | 1.000000 |
| 2   | 1.000000 |
| 4   | 0.999505 |
| 8   | 0.997449 |
| 16  | 1.000000 |
| 32  | 0.968750 |

##### AMI

```{r}
PlusorMinus2_overall(final_AMI_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_AMI_RF_ConMat)
```

| **MIC isolate category** | **Proportion_of_Isolates** |
|:-------------------------|:---------------------------|

|     |           |
|-----|-----------|
| 1   | 0.9473684 |
| 2   | 0.9974874 |
| 4   | 1.0000000 |
| 8   | 0.9375000 |
| 16  | 1.0000000 |

##### AMP

```{r}
PlusorMinus2_overall(final_AMP_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_AMP_RF_ConMat)
```

|     |           |
|-----|-----------|
| 1   | 0.9065156 |
| 2   | 0.9136454 |
| 4   | 0.9098954 |
| 8   | 0.9782609 |
| 16  | 0.7500000 |
| 32  | 0.7861272 |

##### CHL

```{r}
PlusorMinus2_overall(final_CHL_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_CHL_RF_ConMat)
```

| **MIC isolate category** | **Proportion_of_Isolates** |
|:-------------------------|:---------------------------|

|     |           |
|-----|-----------|
| 4   | 0.9400826 |
| 8   | 1.0000000 |
| 16  | 1.0000000 |
| 32  | 0.9589041 |

##### CIP

```{r}
table(CIP.RF.df$CIP)
```

+/-2 category accuracy could not be calculated because there are only 3 levels, so any prediction value is within +/-2 categories

##### COT

```{r}
PlusorMinus2_overall(final_COT_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_COT_RF_ConMat)
```

| **MIC isolate category** | **Proportion_of_Isolates** |
|:-------------------------|:---------------------------|

|       |           |
|-------|-----------|
| 0.125 | 0.9911765 |
| 0.5   | 1.0000000 |
| 2     | 1.0000000 |
| 4     | 0.9250000 |

##### FIS

```{r}
PlusorMinus2_overall(final_FIS_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_FIS_RF_ConMat)
```

| **MIC isolate category** | **Proportion_of_Isolates** |
|:-------------------------|:---------------------------|

|     |           |
|-----|-----------|
| 16  | 0.8963391 |
| 32  | 1.0000000 |
| 128 | 1.0000000 |
| 256 | 0.9498681 |

##### FOX

```{r}
PlusorMinus2_overall(final_FOX_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_FOX_RF_ConMat)
```

| **MIC isolate category** | **Proportion_of_Isolates** |
|:-------------------------|:---------------------------|

|     |           |
|-----|-----------|
| 2   | 0.9352708 |
| 4   | 0.9950690 |
| 8   | 1.0000000 |
| 16  | 0.9200000 |
| 32  | 0.9230769 |

##### TET

```{r}
PlusorMinus2_overall(final_TET_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_TET_RF_ConMat)
```

|     |           |
|-----|-----------|
| 4   | 0.9544271 |
| 8   | 1.0000000 |
| 16  | 1.0000000 |
| 32  | 0.6798088 |

##### TIO

```{r}
PlusorMinus2_overall(final_TIO_RF_ConMat)
```

```{r}
PlusorMinus2_ConMat_Accuracy(final_TIO_RF_ConMat)
```

|       |           |
|-------|-----------|
| 0.125 | 1.0000000 |
| 0.25  | 0.9885877 |
| 0.5   | 0.9976825 |
| 1     | 1.0000000 |
| 2     | 1.0000000 |
| 4     | 1.0000000 |
| 8     | 0.9285714 |
