---
title: "BreakpointCrossing"
format: html
editor: visual
---

## AMC

##### Susceptible/Intermediate Crossing

```{r}
BreakpointCrossing(final_AMC_RF_ConMat, 8)
```

##### Intermediate/Resistant Crossing

```{r}
BreakpointCrossing(final_AMC_RF_ConMat, 16)
```

## AMI

All isolates were susceptible so both are 0%

##### Susceptible/Intermediate Crossing: 0%

##### Intermediate/Resistant Crossing: 0%

## AMP

##### Susceptible/Intermediate Crossing

```{r}
BreakpointCrossing(final_AMP_RF_ConMat, 8)
```

##### Intermediate/Resistant Crossing

```{r}
BreakpointCrossing(final_AMP_RF_ConMat, 16)
```

## CHL

##### Susceptible/Intermediate Crossing

```{r}
BreakpointCrossing(final_CHL_RF_ConMat, 8)
```

##### Intermediate/Resistant Crossing

```{r}
BreakpointCrossing(final_CHL_RF_ConMat, 16)
```

## CIP

There is no intermediate category for ciprofloxacin

##### Susceptible/Resistant Crossing

```{r}
BreakpointCrossing(final_CIP_RF_ConMat, 0.06)
```

## COT

No intermediate category for COT

##### Susceptible/Resistant Crossing

```{r}
BreakpointCrossing(final_COT_RF_ConMat, 2)
```

## FIS

All isolates were susceptible so both are 0%

##### Susceptible/Intermediate Crossing

##### Intermediate/Resistant Crossing

## FOX

##### Susceptible/Intermediate Crossing

```{r}
BreakpointCrossing(final_FOX_RF_ConMat, 8)
```

##### Intermediate/Resistant Crossing

```{r}
BreakpointCrossing(final_FOX_RF_ConMat, 16)
```

## TET

##### Susceptible/Intermediate Crossing

```{r}
BreakpointCrossing(final_TET_RF_ConMat, 4)
```

##### Intermediate/Resistant Crossing

```{r}
BreakpointCrossing(final_TET_RF_ConMat, 8)
```

## TIO

##### Susceptible/Intermediate Crossing

```{r}
BreakpointCrossing(final_TIO_RF_ConMat, 2)
```

##### Intermediate/Resistant Crossing

```{r}
BreakpointCrossing(final_TIO_RF_ConMat, 4)
```
